---
title: "Preliminary analysis of sweetpotato yield loss data"
author: "Lucky Mehra"
date: "9/18/2020"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

```{r load-packages}
if(!require(pacman)){
  install.packages("pacman")
}

pacman::p_load(tidyverse,
               usdata,
               janitor,
               fs,
               gt,
               plotrix)


yieldloss <- read_csv(here::here('data/clean-data', 'yieldloss.csv'))
```

The dataset contains data from `r nrow(yieldloss)` studies from sweetpotato growing
states of the country.

## Study count by different categories

Study count by **Principal Investigator**.

```{r}
yieldloss %>% 
  count(pi, sort = TRUE) %>% 
  gt() %>% 
  cols_label(
    pi = "Principal Investigator"
  )
```

Study count by **sweetpotato variety**.

```{r}
yieldloss %>% 
  count(variety, sort = TRUE) %>% 
  gt() %>% 
  cols_label(
    variety = "Sweetpotato variety"
  )
```

Study count by **soil type**.

```{r}
yieldloss %>% 
  count(soil_type, sort = TRUE) %>% 
  gt() %>% 
  cols_label(
    soil_type = "Soil type"
  )
```

Study count by **primary weed type**.

```{r}
yieldloss %>% 
  count(primary_weed, sort = TRUE) %>% 
  gt() %>% 
  cols_label(
    primary_weed = "Primary weed"
  )
```

Study count by **irrigation type**.

```{r}
yieldloss %>% 
  count(irrigation_type, sort = TRUE) %>% 
  gt() %>% 
  cols_label(
    irrigation_type = "Type of irrigation"
  )
```

Study count by **location**.

```{r}
yieldloss %>% 
  count(location, sort = TRUE) %>% 
  gt() %>% 
  cols_label(
    location = "Location"
  )
```

## Figures

Distribution of marketable yield loss across `r nrow(yieldloss)` studies.

```{r figures-histogram}
# histogram  of marketable loss
hist_market <- yieldloss %>% 
  ggplot(mapping = aes(x = marketable_loss)) +
  geom_histogram() +
  ggtitle("Distibution of marketable yield loss") +
  ylab("Number of studies")

hist_market

# histogram of total loss
# hist_total <- yieldloss %>% 
#   ggplot(mapping = aes(x = total_loss)) +
#   geom_histogram()
# 
# hist_market / hist_total

# graphs of marketable yield loss and total yield loss are very similar
# therefore, only goint to plot marketable yield loss
```

```{r, include=FALSE}
# import the custom function to make plots
source(here::here("R", "functions.R"),
       local = knitr::knit_global())

```

Average marketable loss by **year**.

```{r}
# year
plot_market_loss_by(year)
```

Average marketable loss by **location**.

```{r}
# location
plot_market_loss_by(location)
```

Average marketable loss by **variety**.

```{r}
# variety
plot_market_loss_by(variety)
```

Average marketable loss by **primary weed**.

```{r}
# primary weed
plot_market_loss_by(primary_weed)
```

Average marketable loss by **soil type**.

```{r}
# soil type
plot_market_loss_by(soil_type)
```

Average marketable loss by **irrigation type**.

```{r}
# irrigation type
plot_market_loss_by(irrigation_type)
```

## Tables

Create a table to report average marketable yield loss by state.

```{r loss-by-state}
loss_by_state <- yieldloss %>% 
  separate(col = location,
           into = c("city", "state"),
           sep = "_") %>% 
  mutate(state = abbr2state(state),
         state = if_else(state %in% c("Louisiana",
                                       "Mississippi",
                                       "North Carolina"), state, "Other States")) %>% 
  group_by(state) %>% 
  summarise(marketable_loss = mean(marketable_loss, na.rm = TRUE),
            number_of_studies = n())

loss_by_state %>% 
  arrange(desc(number_of_studies)) %>% 
  mutate_if(is.numeric, round, digits=1) %>% 
  gt()
```

<br>

<br>

Merge the above table with sweetpotato production and area harvested data from USDA
NASS.

```{r}
data_dir <- here::here('data/raw-data/usda-nass')

csv_files <- fs::dir_ls(data_dir, regexp = "\\.csv$")

usda_nass <- csv_files %>% 
  map_dfr(read_csv) %>% 
  remove_empty() %>% 
  clean_names()

usda_nass_1 <- usda_nass %>% 
  dplyr::filter(reference_period == 'MARKETING YEAR') %>% 
  mutate(reference_period = "YEAR") %>% 
  dplyr::select(location, price_received_in_cwt) %>% 
  left_join(usda_nass %>% 
              dplyr::select(-price_received_in_cwt) %>% 
              dplyr::filter(reference_period == 'YEAR'), by = "location") %>% 
  mutate(state = str_to_title(location)) %>% 
  dplyr::select(state, area_harvested_in_acres,
                production_in, # production in dollars
                production_in_cwt,
                yield_in_cwt_acre,
                price_received_in_cwt
                )

potential_loss <- loss_by_state %>% 
  left_join(usda_nass_1) %>% 
  rename(production_in_dollars = production_in) %>% 
  mutate(loss_in_dollars = (production_in_dollars * marketable_loss)/100,
         loss_in_cwt = (production_in_cwt * marketable_loss)/100,
         loss_in_yield_cwt_per_acre = (yield_in_cwt_acre * marketable_loss)/100)

potential_loss %>% 
  relocate(state, marketable_loss, loss_in_dollars,
           loss_in_cwt, loss_in_yield_cwt_per_acre) %>% 
  mutate_if(is.numeric, round, digits=1) %>% 
  dplyr::select(-loss_in_yield_cwt_per_acre,
                -price_received_in_cwt
                  ) %>% 
  gt() %>% 
  cols_label(
    state = "State",
    marketable_loss = "Marketable loss (%)",
    loss_in_dollars = "Loss in $s",
    loss_in_cwt = "Loss in cwt",
    number_of_studies = "Number of studies",
    area_harvested_in_acres = "Harvested (acres)",
    production_in_cwt = "Production in cwt",
    production_in_dollars = "Production in $s",
    yield_in_cwt_acre = "Yield (cwt/acre)"
  ) %>% 
    tab_footnote(
    footnote = "One cwt = 100 lbs",
    locations = cells_column_labels(
      columns = vars(loss_in_cwt,
                     production_in_cwt,
                     yield_in_cwt_acre))
  )

```


